---
title: "RMD work pack"
author: "Perfect Presenter"
date: '2022-06-14'
output: html_document
---

```{r map graphics}
if (!require("usmap", quietly = TRUE)) install.packages("usmap")
install.packages("BiocManager")

library(usmap)
library(stringr)
library(ggplot2)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if("e1071" %in% rownames(installed.packages()) == FALSE) {
  install.packages("e1071")}
if("caTools" %in% rownames(installed.packages()) == FALSE) {
  install.packages("caTools")}
if("caret" %in% rownames(installed.packages()) == FALSE) {
  install.packages("caret")}
  if("moments" %in% rownames(installed.packages()) == FALSE) {
  install.packages("moments")}
    
library(readr)
library(dplyr)
library(tidyr)
require(e1071)
require(caTools)
require(caret)
library(class)

```

## Basics 
### Input, group_by, sum, change column names, merge

1.	How many breweries are present in each state?
2.	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)

```{r}
Breweries <- read_csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv")
Beers <- read_csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Beers.csv", col_types = cols(Beer_ID = col_integer(), 
    Brewery_id = col_integer()), na = "NA")

# rename columns
rename(Beers, Brew_ID = Brewery_id,Beer=Name) -> Beers
rename(Breweries, state = State,Brewery=Name) -> Breweries

# group by state, count breweries in state, order by count and state.
Q1 <- Breweries %>% group_by(state) %>% summarise(Density = n()) %>% arrange(desc(Density),state)

# merge many to one 
Q2 <- inner_join(Beers,Breweries,by ="Brew_ID")

head(Q2)
tail(Q2)

write_csv(Q1,file="StateBrew.csv")
```

one can use Excel to format the data into a 6 column table for easy display

## Choropleth

> Data is dataframe and must have state code and numerical value
size = line size, values is column with data to use for colors
name = name of label and label defines type of label. 
labs = label or title.

```{r}
plot_usmap(regions ="state", data = Q1, size=.2, values = "Density") + 
    scale_fill_continuous( low = "pink", high = "darkred", 
                         name = "Breweries", label = scales::comma) + 
  theme(legend.position = "right") +
  labs(title = "Brewery Density by State") 
#+   theme(panel.background = element_rect(color = "grey90", fill = "lightblue")) 
```




### Calculating medians, using group by , dealing with N/A  

```{r}
medianABV = median(Q2$ABV)
medianABV

medianABVna = median(Q2$ABV, na.rm = TRUE)
medianABVna

Q2 %>% group_by(state) %>% 
  summarize(medianstateABV = median(ABV,na.rm=TRUE),
            maxstateABV = max(ABV, na.rm=TRUE)) -> StateABV
StateABV

# sorting
Q2 %>% arrange(desc(IBU))-> Temp 
head(Temp,1)

# barcharts (use reorder to change the order of the state (as factor) to match maxstateABV)
d <- ggplot(StateABV, aes(y=maxstateABV,x=reorder(as.factor(state),maxstateABV)))
d <- d + geom_col()+ coord_flip() +
  theme(axis.text = element_text(size = 6)) 
d
```

#### scatter plots

```{r}
plot(StateABV$maxstateABV,StateABV$medianstateABV, col = as.factor(StateABV$state))
```

### filtering using text strings. 

>**grepl()**: grepl() function will is used to return the value TRUE if the specified string pattern is found in the vector and FALSE if it is not found.
Syntax: grepl(pattern, string, ignore.case=FALSE)

```{r}
list = grepl("ale",Q2$Style, ignore.case=TRUE)
alelist <- Q2 %>% filter(grepl("ale",Style, ignore.case=TRUE) & 
                           !(grepl("IPA",Style)|grepl("IPA",Beer)))
IPAlist <- Q2 %>% filter((grepl("IPA",Style)|grepl("IPA",Beer)))
alelist$type <- "ale"
IPAlist$type <- "IPA"
Beerlist <- rbind(alelist,IPAlist)
KNND <- as.factor(Beerlist$type)
KNNI <- Beerlist[,c(3,4)]
```

### Create datasets for KNN analysis (one data.frame with outcomes, a
second data.frame with the independent variables (ABV and IBU))



Beers[Beers$ABV>0.12&!is.na(Beers$ABV),]
Q2[grepl("Bud",Q2$Beer,ignore.case=TRUE),]

